---
title: "2212064"
subtitle: "MA335 Lab Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(faraway)
```



```{r, echo=TRUE, warning=FALSE}
#1(a)
setwd("m:/pc/downloads")
#1(b)
test <- read.csv("Real estate.csv",header=T)
attach(test)
head(test)
```


```{r, echo=TRUE, warning=FALSE}
#1(c)
price
summary(price)
```

```{r, echo=TRUE, warning=FALSE}
#1(d)
dim(test)
```
From the below summary we can say that y = 46.53 -0.0072733x which means that for every decrease in the value of x the distance to the station increases by 46.53m. The value of P is  <2e-16 hence we can even reject null hypothesis. The R2 is 51.52% and adjusted R2 is 51.28 with a slight decrease.Hence we can that the predictor is s statistically significant at the 1% level.


```{r, echo=TRUE, warning=FALSE}
#2(a)
model1 <- lm(price~MRT.station)
summary(model1)
```


```{r, echo=TRUE, warning=FALSE}
#2(b)
plot(MRT.station,price, xlab = "Distance", ylab = "Price", main = "Scatter Plot")
abline(model1, lwd=3, col="blue")
```
## 3
From the below analysis, we see from the p-value which is smaller than < 2.2e − 16 that the null
hypothesis is rejected, therefore we have evidence that model2 fits the data better. This is confirmed when
we check the Adjusted R-squared which is 0.5128 for model1 and increases to 73.58% for model2. Also the
Residual standard error from 9.431 for model1 reduces to 6.945 for model2. The F-statistics which decreases
from 210.4 to 80.17 confirms the above conclusion, that model2 is better than model1.
```{r, echo=TRUE, warning=FALSE}
#3
model_2 <- lm(price~date+age+MRT.station+stores+latitude+longitude+properties, data=test)
model_2
```

```{r, echo=TRUE, warning=FALSE}
#3
summary(model_2)

```

```{r, echo=TRUE, warning=FALSE}
#3
summary(model1)

```

```{r, echo=TRUE, warning=FALSE}
#3
anova(model1, model_2)
```
##4

Below is the correlation between various columns

```{r, echo=TRUE}
#4

cor(test)

```


```{r, echo=TRUE}
#5

aov(price~as.factor(price))

```











